---
title: "Developing Data Products Pitch"
author: "Daniel De Leonardis"
date: "7 September 2017"
output: ioslides_presentation
---

<style type="text/css">
.reduced {
      font-size: 20px;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE)
```

## Australian General Record of Incidence of Mortality (GRIM) Overview

<div class="reduced">
The GRIM books house national level, historical and recent deaths data for specific causes of death. The tables present age- and sex-specific counts and rates of deaths, and age-standardised death rates, for all causes and for International Statistical Classification of Diseases and Related Health Problems.

The data goes back to 1907, but data may not be available for before a certain year for a particular cause of death. In general most causes of death apply to both sexes but there are instances where it may apply to only sex e.g. prostrate cancer for males.

A subset of this data is taken to enable to plot the number of deaths for a selected cause of death by year and sex as well as the total for all sexes.

Further Information and the data set can be found at the [Australian Government Website](http://data.gov.au/dataset/grim-books).
</div>


## Subsetting the GRIM Data

<div class="reduced">
The full data set contains around 300,000.  A short script was written to reduce the size of the data set to around 10,000 records (primarily to substantially reduce the load time in the shiny app).

Some cleanup is also performed for the following:

- reduce the size of the cause of death field by removing some unnecessary data with brackets at the end of each row.
- standardise the field names output to camel case.
- Change person to total in the sex field for use in the plot.

Here is a small sample of the subsetted data:
</div>

```{r loadData, message=FALSE, warning=FALSE, cache=TRUE}
library(dplyr)

fileUrl <- url("http://data.gov.au/dataset/488ef6d4-c763-4b24-b8fb-9c15b67ece19/resource/edcbc14c-ba7c-44ae-9d4f-2622ad3fafe0/download/grimdatagovau.csv")
deaths <- read.csv(fileUrl, stringsAsFactors = FALSE)

deaths$cause_of_death <- trimws(gsub( " *\\(.*?\\) *", "", deaths$cause_of_death))

deathTotals <- filter(deaths, AGE_GROUP == "Total") %>%
    na.omit() %>%
    rename(year = YEAR, causeOfDeath = cause_of_death, sex = SEX)  %>%
    select(year, causeOfDeath, sex, deaths) 

deathTotals$sex <- ifelse(deathTotals$sex == "Persons", "Total", deathTotals$sex)
head(deathTotals[4:6, ])

#write.csv(x=deathTotals, file = "./deathTotals.csv", row.names = FALSE)
```

<div class="reduced">
This data set can be found on [GitHub](https://raw.github.com/DanielDeLeonardis/DevelopingDataProductsAssignment/gh-pages/deathTotals.csv).
</div>

## Plotting the Data

<div class="reduced">
The shiny app allows the user to select from a list of causes of death to display a particular plot. The data set is filtered by the cause of death and then Plotly is used to display the plot. An example of a plot is shown here:
</div>

```{r plotData, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
library(plotly)

filterByDeathCause <- function(df, deathCause) {
    dplyr::filter(df, causeOfDeath %in% deathCause) %>%
        select(year, sex, deaths)
}

plotterBySex <- function(year, deaths, sex) {
    lTitle = paste("Australian Deaths", min(year), "to", max(year))
    plot_ly(x = year,
            y = deaths,
            color = sex,
            type = "scatter",
            mode = "lines+markers") %>%
        layout(title = lTitle,
               xaxis = list(title = "Year"),
               yaxis = list(title = "Deaths"))
}

deathsByCause <- filterByDeathCause(df=deathTotals, deathCause="Heart failure")
plotterBySex(year = deathsByCause$year,
             deaths = deathsByCause$deaths,
             sex = deathsByCause$sex)
```

## Code and Shiny Application

<div class="reduced">

The code for the following:

- this Presentation
- subset the GRIM data set
- Shiny application for the interactive plotting of data by cause of death

can be found here:

- https://github.com/DanielDeLeonardis/DevelopingDataProductsAssignment


The Shiny applications for the interactive plotting of data by cause of death can be found here: 

- https://danieldeleonardis.shinyapps.io/developing_data_products/

Note: This presentation was created, as suggested on stackoverflow, using R Markdown (ioslides_presentation) rather than an R Presentation as plotly does not play well with it e.g. see the following post:

- https://stackoverflow.com/questions/42045969/error-in-filecon-rb-cannot-open-the-connection
</div>

